FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2018-06-11 ECOOP18/sections/Auxiliary.tex
spellcheck

\subsection{Finding the Most Specific Origin: \mostSpecific}\label{sec:mostSpecific}
We proceed to give the definitions of two core functions that support method lookup, namely \mostSpecific{} and \mostSpecificOverride. Generally,
$\mostSpecific(m, I, J)$ finds the set of most specific interfaces where $m$ is originally defined. Interfaces in this set should be above interface $I$ and along path $J$. Finally with $\prune$ (defined in Section~\ref{sec:otherdefs})
the overridden interfaces will be filtered out.

@@

$\canInstantiate(I)$ checks whether interface $I$ can be instantiated by the keyword $\kwnew$.
$\mostSpecific(m, I, I)$ represents the set of branches $I$ inherits on method $m$. $I$ can be instantiated
if and only if for every branch, the most specific implementation is non-abstract. \yanlin{removed 'unambiguous'.}
%%%%============================ canInstantiate end ================%%%%%%%%

>>>>

\subsection{Finding the Most Specific Origin: \mostSpecific}\label{sec:mostSpecific}
We proceed to give the definitions of two core functions that support method lookup, namely, \mostSpecific{} and \mostSpecificOverride. Generally,
$\mostSpecific(m, I, J)$ finds the set of most specific interfaces where $m$ is originally defined. Interfaces in this set should be above interface $I$ and along path $J$. Finally with $\prune$ (defined in Section~\ref{sec:otherdefs})
the overridden interfaces will be filtered out.

@@

$\canInstantiate(I)$ checks whether interface $I$ can be instantiated by the keyword $\kwnew$.
$\mostSpecific(m, I, I)$ represents the set of branches that $I$ inherits on method $m$. $I$ can be instantiated
if and only if for every branch, the most specific implementation is non-abstract. \yanlin{removed 'unambiguous'.}
%%%%============================ canInstantiate end ================%%%%%%%%

### 2018-06-11 ECOOP18/sections/Formalization.tex
spellcheck

\subsubsection{Expressions \& Values}
Expressions can be standard constructs such as variables, method
invocation, object creation, together with cast expressions. 
Object creation is represented by $\new I$\footnote{In Java the corresponding syntax is $\new I\{\}$.}. Fields and primitive types are not modelled in \MIM{}. 
The casts are merely safe upcasts, and in fact, they can be viewed as
annotated expressions, where the annotation indicates its static type.
The coexistence of static and dynamic types is the key to hierarchical dispatch.
A value
``$(I)\new{J}$''
is the final result of multiple reduction steps evaluating an
expression.

@@

\textsc{(T-Method)} is more interesting since a method can either be an original method or a hierarchical overriding, though
they share the same syntax and method typing rule. $\mostSpecific(m, I, J)$ is a fundamental function,
used to find ``the most specific interfaces that are above $I$ and
along path $J$, and originally defines $m$'' (see
Section~\ref{sec:auxdefs} for full definition).
By ``most specific interfaces'',
it implies that the inherited supertypes are excluded. Thus the condition $\mostSpecific(m, I, J) = \{J\}$ indicates a characteristic of a hierarchical overriding: it must override an original method; the overriding is direct and there does not exist any other original method $m$ in between.
Then $\mbody(m, J, J)$ provides the type of the original method, so hierarchical overriding has to preserve the type. Finally the return expression
is type-checked to be subtype of the declared return type. For the definition of an original method, $I$ equals $J$ and the rule is straightforward. \textsc{(T-AbsMethod)} is a similar rule but works on abstract method declarations.

\textsc{(T-Intf)} defines the typing rule on interfaces. The first condition is obvious, namely its methods need to be well checked. The third
condition checks whether the overriding between original methods preserves typing. In this condition we again use some helper functions defined in  Section~\ref{sec:auxdefs}. $I[m\ \kwoverride\ I]$ is defined if $I$ originally defines $m$, and $\canOverride(m, I, J)$ checks whether $I.m$ has the same type as $J.m$. Generally the preservation of method type is required for any supertype $J$ and any method $m$.

>>>>

\subsubsection{Expressions \& Values}
Expressions can be standard constructs such as variables, method
invocation, object creation, together with cast expressions. 
Object creation is represented by $\new I$\footnote{In Java the corresponding syntax is $\new I\{\}$.}. Fields and primitive types are not modelled in \MIM{}. 
The casts are merely safe upcasts, and in fact, they can be viewed as
annotated expressions, where the annotation indicates its static type.
The coexistence of static and dynamic types is the key to hierarchical dispatch.
A value
``$(I)\new{J}$''
is the final result of multiple reduction steps for evaluating an
expression.

@@

\textsc{(T-Method)} is more interesting since a method can either be an original method or a hierarchical overriding, though
they share the same syntax and method typing rule. $\mostSpecific(m, I, J)$ is a fundamental function,
used to find ``the most specific interfaces that are above $I$ and
along path $J$, and originally defines $m$'' (see
Section~\ref{sec:auxdefs} for full definition).
By ``most specific interfaces'',
it implies that the inherited supertypes are excluded. Thus the condition $\mostSpecific(m, I, J) = \{J\}$ indicates a characteristic of a hierarchical overriding: it must override an original method; the overriding is direct and there does not exist any other original method $m$ in between.
Then $\mbody(m, J, J)$ provides the type of the original method, so hierarchical overriding has to preserve the type. Finally the return expression
is type-checked to be a subtype of the declared return type. For the definition of an original method, $I$ equals $J$ and the rule is straightforward. \textsc{(T-AbsMethod)} is a similar rule but works on abstract method declarations.

\textsc{(T-Intf)} defines the typing rule on interfaces. The first condition is obvious, namely, its methods need to be well checked. The third
condition checks whether the overriding between original methods preserves typing. In this condition we again use some helper functions defined in  Section~\ref{sec:auxdefs}. $I[m\ \kwoverride\ I]$ is defined if $I$ originally defines $m$, and $\canOverride(m, I, J)$ checks whether $I.m$ has the same type as $J.m$. Generally the preservation of method type is required for any supertype $J$ and any method $m$.

### 2018-06-11 ECOOP18/sections/Overview.tex
spellcheck

\section{A Running Example: Drawable Deck}\label{sec:overview}
This section illustrates the problem of unintentional method conflicts,
together with the features of our model for addressing this issue, by
a simple running example. In the following text we will introduce three
problems one by one and have a discussion on possible workarounds and
our solutions.
Problems 1 and 2 are related to hierarchical dispatching, and in 
C++ it is possible to have similar solutions to both
problems. Hence it is important to emphasize that, with respect to
hierarchical dispatching, our model is not a novel
mechanism. Instead, 
inspired by the C++ solutions, our contribution is formalizing a minimal calculus
of this feature together with a proof of type soundness. However, for the final problem, there is no satisfactory approach
in existing languages, thus what we propose is a novel feature (hierarchical
overriding) with the corresponding formalization of that feature.

@@

\subsection{Problem 1: Basic Unintentional Method Conflicts}\label{subsec:problem1}
Suppose that two components \lstinline|Deck| and \lstinline|Drawable| 
have been developed in a system. \lstinline|Deck| represents a deck
of cards and defines a method \lstinline|draw| for drawing a card from the
deck.  \lstinline|Drawable| is an interface for graphics that
can be drawn and also includes a method called \lstinline|draw| for
visual display. For simple illustration, the default implementation of
\lstinline|draw| in \lstinline|Drawable| only creates a blank canvas
on the screen, while the \lstinline|draw| method in \lstinline|Deck| simply
prints out a message \lstinline|"Draw a card."|.

@@

  \item \textit{II. Refactor \lstinline|Drawable| and/or \lstinline|Deck| to rename the methods.} If
  the source code for \lstinline|Drawable| or \lstinline|Deck| is available
  then it may be possible to rename one of the \lstinline|draw|
  methods. However this approach is non-modular, as it requires 
  modifying existing code and becomes impossible if the code is unavailable.

@@

Note that, although the example here only shows one conflicting method
being overriden, hierarchical overriding allows (as expected) multiple conflicting
methods to be overriden in the same class.

>>>>

\section{A Running Example: Drawable Deck}\label{sec:overview}
This section illustrates the problem of unintentional method conflicts,
together with the features of our model for addressing this issue, by
a simple running example. In the following text, we will introduce three
problems one by one and have a discussion on possible workarounds and
our solutions.
Problems 1 and 2 are related to hierarchical dispatching, and in 
C++ it is possible to have similar solutions to both
problems. Hence it is important to emphasize that, with respect to
hierarchical dispatching, our model is not a novel
mechanism. Instead, 
inspired by the C++ solutions, our contribution is formalizing a minimal calculus
of this feature together with a proof of type soundness. However, for the final problem, there is no satisfactory approach
in existing languages, thus what we propose is a novel feature (hierarchical
overriding) with the corresponding formalization of that feature.

@@

\subsection{Problem 1: Basic Unintentional Method Conflicts}\label{subsec:problem1}
Suppose that two components \lstinline|Deck| and \lstinline|Drawable| 
have been developed in a system. \lstinline|Deck| represents a deck
of cards and defines a method \lstinline|draw| for drawing a card from the
deck.  \lstinline|Drawable| is an interface for graphics that
can be drawn and also includes a method called \lstinline|draw| for
visual display. For simple illustration, the default implementation of the 
\lstinline|draw| in \lstinline|Drawable| only creates a blank canvas
on the screen, while the \lstinline|draw| method in \lstinline|Deck| simply
prints out a message \lstinline|"Draw a card."|.

@@

  \item \textit{II. Refactor \lstinline|Drawable| and/or \lstinline|Deck| to rename the methods.} If
  the source code for \lstinline|Drawable| or \lstinline|Deck| is available
  then it may be possible to rename one of the \lstinline|draw|
  methods. However, this approach is non-modular, as it requires 
  modifying existing code and becomes impossible if the code is unavailable.

@@

Note that, although the example here only shows one conflicting method
being overridden, hierarchical overriding allows (as expected) multiple conflicting
methods to be overridden in the same class.
